---
title: "Project"
author: "Belinda Hu"
date: "4/3/2020"
output: html_document
---
# General topic

My general research topic is to understand the factors that contribute to a startup being founded and its success. Below is the preliminary work I have done.

Github repo: https://github.com/belindahu/startups 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(scales)
library(sp)
library(janitor)
library(sf)
library(tidycensus)
library(tigris)

census_api_key("03bf7eeb14e29c5d6fd4073fce357ba84f0017cc")

```


```{r load_startups, echo=FALSE}

s_offices <- read_csv("raw-data/startup-investments/offices.csv", 
                      col_types = cols(
  id = col_double(),
  object_id = col_character(),
  office_id = col_double(),
  description = col_character(),
  region = col_character(),
  address1 = col_character(),
  address2 = col_character(),
  city = col_character(),
  zip_code = col_character(),
  state_code = col_character(),
  country_code = col_character(),
  latitude = col_double(),
  longitude = col_double(),
  created_at = col_logical(),
  updated_at = col_logical()
)) %>% 
  select(id, object_id, office_id, region, city, zip_code, state_code, country_code)

s_objects <- read_csv("raw-data/startup-investments/objects.csv",
                      col_types = cols(
  .default = col_character(),
  entity_id = col_double(),
  parent_id = col_logical(),
  founded_at = col_date(format = ""),
  closed_at = col_date(format = ""),
  logo_width = col_double(),
  logo_height = col_double(),
  first_investment_at = col_date(format = ""),
  last_investment_at = col_date(format = ""),
  investment_rounds = col_double(),
  invested_companies = col_double(),
  first_funding_at = col_date(format = ""),
  last_funding_at = col_date(format = ""),
  funding_rounds = col_double(),
  funding_total_usd = col_double(),
  first_milestone_at = col_date(format = ""),
  last_milestone_at = col_date(format = ""),
  milestones = col_double(),
  relationships = col_double(),
  created_at = col_datetime(format = ""),
  updated_at = col_datetime(format = "")
                      )) 

```

```{r load_census, echo=FALSE, message = FALSE}

zip_county_cross <- read_xlsx("raw-data/ZIP_COUNTY_122014.xlsx") %>% 
  select(ZIP, COUNTY)

income1 <- read_csv("raw-data/2014_Income.csv", skip = 1,
                   col_types = cols(
  Id2 = col_character(),
  Id = col_character(),
  "Median income (dollars); Estimate; Households" = col_character()
)) %>%
  rename(median_income = "Median income (dollars); Estimate; Households", county = Geography) %>% 
  select(Id2, county, median_income)


```

```{r join_data, echo = FALSE}

# need to remove rows without a county

startupbycounty <- s_offices %>% 
  filter(country_code == "USA") %>% 
  full_join(zip_county_cross, by = c("zip_code" = "ZIP")) 

# join zip_county_cross with income dataset 

income2 <- income1 %>% 
  full_join(zip_county_cross, by = c("Id2" = "COUNTY")) %>% 
  rename(GEOID = Id2)

# join income data with office data

incomebyoffice1 <- s_offices %>% 
  left_join(income2, by = c("zip_code" = "ZIP")) %>% 
  select(id, office_id, region, city, zip_code, state_code, country_code, county, median_income) %>% 
  filter(country_code == "USA")
```

# Visualization Work

```{r CA income, echo = FALSE}

# visualize median income of Calfornia by county

# california <- incomebyoffice %>% 
#   filter(state_code == "CA") %>% 
#   ggplot(aes()) + 
#   geom_bar()
# 
# california

map_income <- get_acs(geography = "state",
                 variables = "B19013_001",
                 year = 2018,
                 geometry = TRUE) %>% 
  rename(state = NAME) 

# ggplot(map_income, aes(fill = estimate)) + 
#   geom_sf()
  

# need to join map_income to incomebyoffices to have geometry variable

# incomebyofficeCA <- incomebyoffice1 %>% 
#   left_join(map_income, by = c("county" = "county")) %>% 
#   filter(state_code == "CA")
# 
# ggplot(incomebyofficeCA, aes(geometry = geometry, fill = median_income)) +
#   geom_sf() 

```

```{r number of startups in each state BAR, echo = FALSE}

bystate <- s_offices %>% 
  count(state_code) %>% 
  na.omit() 

plot <- bystate %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(., aes(x = reorder(state_code, -n), y = n)) +
  geom_bar(stat = "identity") + 
  labs(title = "Top Ten States with the Most Startups", 
       subtitle = "By Number of Offices Per State", 
       x = "State", 
       y = "Number of Startups") +
  theme_classic() +
  geom_text(aes(label = comma(n, accuracy = 1), 
                vjust = -0.2)
  )

ggsave("topten_plot.png", plot = plot, path = "startups_shiny")


```

```{r number of startups per state MAP, echo = FALSE}

us <- states(year = 2018)
plot(us)

us_states <- read_csv("raw-data/us_states.csv",
                      col_types = cols(
  state_name = col_character(),
  state_code = col_character()
))

bystate_mapping <- bystate %>% 
  full_join(us_states, by = c("state_code" = "state_code")) %>% 
  full_join(map_income, by = c("state_name" = "state")) %>% 
  select(-variable, -estimate, -moe, -GEOID)
  
ggplot(bystate_mapping, aes(fill = n, geometry = geometry)) +
  geom_sf()

```

Many of these are commented out as this is still a work in progress.

## Action items:
- configure Git LFS to upload objects.csv file
- brainstorm more potential joins between datasets (as the startup datasets has multiple usable csv files)
- because these datasets are large and messy, I need to clean up the data processing to remove any warnings
- clean up s_objects

## Potential directions:
- visualize types of startups by state
- correlate median income to number of startups
- visualize where founders of startups are from
- census shapefiles
- animate funding rounds over time
